Pre-processing high frequency animal tracking data

### 2020-08-13 01_background.Rmd
remove minor sections

# Background

Tracking individual movement has become a firmly established component of animal ecology, helped by the adoption of the movement ecology paradigm (Nathan et al. 2008; Holyoak et al. 2008) and the rapid proliferation of animal tracking technologies (Kays et al. 2015; Hussey et al. 2015). The animal tracking revolution has offered unprecedented insights into the causes and consequences of animal space-use, from the mechanisms behind collective behaviour (e.g. Strandburg-Peshkin et al. 2015), to how animal migrations engineer landscape level processes (Geremia et al. 2019).

## Install `atlastools`

```{r install_package}
# use either devtools or remotes to install
# install.packages("devtools")

# install using devtools
devtools::install_github("pratikunterwegs/atlastools")
```

>>>>

## Installing `atlastools`

This paper refers extensively to the `R` package `atlastools` (Gupte 2020), which can be installed from an online repository.

```{r install_package, eval=FALSE}
# use either devtools or remotes to install
install.packages("devtools")

# installation using devtools
devtools::install_github("pratikunterwegs/atlastools")
```

### 2020-08-13 02_filtering_on_spatio_temporal_bounds.Rmd
remove minor sections

# save figure
ggsave("figures/fig_filter_bbox.png", dpi = 300, width = 8, height = 2)
```

@@

Explicit spatial filtering is a finer method which examines whether the tracking data lie within a polygon, which may be irregular. This typically involves converting the data’s position coordinates to a multi-point spatial geometry, and finding the intersection with a polygon or multi-polygon geometry representing a study area. 
Combined with knowledge of species’ biology and the study site, explicit spatial filtering can be used to remove obviously erroneous positions such as those that would place an individual in an area it is physiologically incapable of traversing. The main drawback to explicit spatial filtering is that it is computationally intensive, and it is advisable to apply a coarse bounding box filter first in order to reduce the computational load, and only then filter by an irregular spatial feature.

>>>>

# save figure
ggsave("figures/fig-filter-bbox.png", dpi = 300, width = 8, height = 2)
```

@@

## Explicit spatial filtering

```{r make_non-bbox_polygon}
# make a polygon buffer around centroid
library(sf)
data <- data.table::fread("data/sim_data_sec_02_01.csv")
centroid <- st_point(c(median(data$x), median(data$y)))
buffer <- st_buffer(centroid, 5)
```

```{r spatial_filter_function}
# an ad-hoc spatial filter function which assumes two coordinate columns
# from a data table or dataframe
do_spatial_filter <- function(data,
                              spatial_object) {

  data_coords <- data[, c("x", "y")]
  data_coords <- st_sfc(st_multipoint(as.matrix(data_coords, byrow = TRUE)))
  data_coords <- st_cast(data_coords, "POINT")
  overlaps <- sf::st_intersects(data_coords, spatial_object)

  # not length but lengths
  data <- data[lengths(overlaps) > 0, ]

  return(data)
}
```

```{r do_filter_spatial}
# filter the data
filtered_data <- do_spatial_filter(data = data, 
                                   spatial_object = buffer)

# assign types and plot
data$type <- "raw"
filtered_data$type <- "filtered"

# bind rows
data <- rbind(data, filtered_data)

# assign levels
data$type <- factor(data$type,
                    levels = c("raw", "filtered"))
```

```{r fig_spatial_filter}
fig_spatial_filter <- 
  ggplot(data)+
  geom_path(aes(x, y),
            col = "blue")+
  theme_test()+
  geom_sf(data = buffer, 
          fill = "grey", 
          alpha = 0.2)+
  facet_grid(~ type)+
  theme(axis.text = element_blank(),
        axis.title = element_blank())+
  labs(x = "X coordinate",
       y = "Y coordinate")

# save figure
ggsave("figures/fig-spatial-filter.png", dpi = 300, width = 5, height = 2)
```

## Bounding box before spatial filter

### Make a larger dataset

Here we simulate a dataset of 50,000 positions, and save it to file for further use.

```{r sim_data_large, eval=FALSE}
# uses smoove from Gurarie et al. (2017)
library(smoove)
nu <- 2
tau <- 5
dt <- .1 
ucvm1 <- simulateUCVM(nu=nu, tau=tau, T.max = 5e4L, dt = dt)

# add a reflection to x_max + x_max / 10, y_max + y_max / 10
data <- as.data.table(ucvm1$XY)
data[, time := seq_len(nrow(data))]

fwrite(data, "data/sim_data_large.csv")
```

### Resample large data

```{r get_resample_prop}
# powers of 10
resample_rate <- c(
  # 1, 2, 3, 10, 
                   100)

# get range of coordinates
data_range <- sapply(data[, c("x", "y")], range)

# get resample parameters
resample_parameters <- CJ(rate = resample_rate,
                          bbox_size = resample_rate)
```

```{r}
do_resample_benchmark <- function(data,
                                  resamp_rate, prop_bbox) {

  bbox <- sapply(data[, c("x", "y")], range) / prop_bbox
  data <- data[floor(sample(nrow(data) / resamp_rate)), ]

  microbenchmark::microbenchmark(
    atl_filter_bounds(data,
                      x_range = bbox[, "x"],
                      y_range = bbox[, "y"],
                      remove_inside = FALSE),
    unit = "s"
  )
}
```

### Benchmarking bounding box filters

```{r benchmark_bbox_filter}
library(microbenchmark)

microbenchmark(
  atlastools::atl_filter_bounds(data,
                                x = "x", y = "y",
                                x_range = c(-15, 0),
                                y_range = c(-35, -15)),
  unit = "s"
)
```

### 2020-08-13 index.rmd
remove minor sections

--- 
knit: "bookdown::render_book"
title: "Pre-processing high frequency animal tracking data"
author: "Pratik R. Gupte and others"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
# bibliography: [refs.bib]
biblio-style: apalike
link-citations: yes
github-repo: pratikunterwegs/atlas-best-practices
description: "TRES Tidyverse Tutorial"
---

>>>>

--- 
knit: "bookdown::render_book"
title: "Pre-processing high frequency animal tracking data"
author: "Pratik R. Gupte and others"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::pdf_book
documentclass: book
# bibliography: [refs.bib]
biblio-style: apalike
link-citations: yes
github-repo: pratikunterwegs/atlas-best-practices
description: "TRES Tidyverse Tutorial"
---
